---
title: "p8105_hw3_qx2199"
author: "Qianhui Xu"
date: '2020-10-08'
output: github_document
---


```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

##Load the Instacart data.

```{r}
data("instacart")
```

The Instacart dataset contains `r nrow(instacart)` rows 
and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- product ID, product name, aisle, department, and some numeric codes. 

##How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```
As we can see, there are 134 aisles, and  " fresh vegetables "are which the most items from.

##Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize the plot.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

The top 2 aisles are "fresh vegetables" and "fresh fruits".

##Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in the table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

*The top 3 items in baking ingredients aisle are "Light Brown Sugar", "Pure Baking Soda", and "Cane Sugar".

*The top 3 items in dog food care aisle are "Snack Sticks Chicken & Rice Recipe Dog Treats", "Organix Chicken & Brown Rice Recipe", and "Small Dog Biscuits".

*The top 3 items in packaged vegetables fruits aisle are "Organic Baby Spinach", "Organic Raspberries", and "Organic Blueberries".

##Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format the table as a 2 x 7 table.

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```


## Problem 2

##Load, tidy, and otherwise wrangle the accelerometers data

```{r}
accel_df = read.csv('./accel_data.csv') %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity_counts"
  ) %>% 
mutate(day = factor (day, levels = c ("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")))%>%

  mutate(day_type = case_when(
      day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
      day %in% c("Saturday", "Sunday")~ "weekend")) %>%

  mutate(minute = as.numeric(minute)) %>%
arrange(week,day)
```
##data discription:
The tidy accelerometers dataset include `r ncol(accel_df)` variables and `r nrow(accel_df)` rows.The variable names are 'week', 'day','minute','activity counts' 'day id' and 'day type'.


##Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals.

```{r}
 
accel_df_new = accel_df %>% 
  group_by(week, day) %>% 
  summarise(total_activity_per_day = sum(activity_counts)) %>% 
  arrange(week, day) %>% 
pivot_wider(
  names_from = day,
  values_from = total_activity_per_day
)
```

*The person tends to be less active on the Saturday in week 4 and week 5. Also, the person tends to have a relative stable activity exercise during the weekdays and have relatively more activity exercise on Fridays. 

##Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.


```{r}
accel_df_new %>% 
  group_by(week, day) %>% 
  ggplot(aes(x = day, y = total_activity_per_day, colour = day)) + geom_point(alpha = 0.5) + geom_smooth (se = FALSE) +
    labs( title = "Daily Activity Plot",
      x = "Day of the week",
      y = "Daily activity count"
    )
```




















